---
title: "Ames Housing Final Project"
date: "Last updated on `r Sys.Date()`"
author: "Group E"
output: 
  html_document:
    fig_width: 8
    fig_height: 6
---




```{r Loading Packages, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(tidyr)
library(dplyr)
library(ggcorrplot)
library(ggthemes) 
library(hrbrthemes)
library(viridis)
library(leaps)
library(boot) 
library(MASS)
library(glmnet)
library(ggpubr)
library(gridExtra)
library(knitr)
library(boot)
library(gam)
library(GGally) 
library(plotly)
library(tidyverse)
library(scales)
options(scipen=10000)
```

```{r loading data, message=TRUE, warning=FALSE, echo = FALSE}
AmesHousing <- read.csv("AmesHousing.csv")
```


```{r Changing column names, message=FALSE, warning=FALSE, echo = FALSE}
#changing column names
colnames(AmesHousing)[3] <- c("home_type")
colnames(AmesHousing)[6] <- c("lot_area")
colnames(AmesHousing)[19] <- c("overall_qual")
colnames(AmesHousing)[21] <- c("year_built")
colnames(AmesHousing)[51] <- c("full_bath_abv_grd")
colnames(AmesHousing)[56] <- c("tot_rms_abv_grd")
colnames(AmesHousing)[82] <- c("saleprice")
```

```{r, echo = FALSE}
#Exploratory Analysis (Shannon's Variables)
#Remove levels with less than 10 observations
# AmesHousing <- AmesHousing[!(AmesHousing$"Garage Qual" =="Ex" | AmesHousing$"Garage Qual" == "Po"),]
# AmesHousing <- AmesHousing[!(AmesHousing$"Exterior 1st" =="AsphShn" | AmesHousing$"Exterior 1st" == "BrkComm"| AmesHousing$"Exterior 1st" == "CBlock"| AmesHousing$"Exterior 1st" == "ImStucc"| AmesHousing$"Exterior 1st" == "PreCast"| AmesHousing$"Exterior 1st" == "Stone"),]
# AmesHousing <- AmesHousing[!(AmesHousing$"Foundation" =="Wood"),]
# AmesHousing <- AmesHousing[!(AmesHousing$"Bsmt Qual" =="Ex" | AmesHousing$"Bsmt Qual" == "Po"),]
# AmesHousing <- AmesHousing[!(AmesHousing$"Heating QC" =="Po"),]
# AmesHousing <- AmesHousing[!(AmesHousing$"Roof Style" =="Shed"),]
# AmesHousing <- AmesHousing[!(AmesHousing$"Foundation" == "Stone"),]
# AmesHousing <- AmesHousing[!(AmesHousing$"Roof Style" == "Mansard"),]

# #All of the features are in character format, so need to convert to factors
# AmesHousing$Garage.Qual <- as.factor(AmesHousing$Garage.Qual)
# AmesHousing$Exterior.1st <- as.factor(AmesHousing$Exterior.1st)
# AmesHousing$Foundation <- as.factor(AmesHousing$Foundation)
# AmesHousing$Bsmt.Qual <- as.factor(AmesHousing$Garage.Qual)
# AmesHousing$Heating.QC <- as.factor(AmesHousing$Heating.QC)
# AmesHousing$Roof.Style <- as.factor(AmesHousing$Roof.Style)
# #Re-code
# AmesHousing$Garage.Qual <- factor(AmesHousing$Garage.Qual, levels = c("Fa", "TA", "Gd"))
# AmesHousing$`Bsmt.Qual` <- factor(AmesHousing$`Bsmt.Qual`, levels = c("Fa", "TA", "Gd"))
# AmesHousing$`Heating.QC` <- factor(AmesHousing$`Heating.QC`, levels = c('Fa', 'TA', 'Gd', 'Ex'))
# #Rename
# levels(AmesHousing$Garage.Qual) <- c("1", "2", "3")
# levels(AmesHousing$Bsmt.Qual) <- c("1", "2", "3")
# levels(AmesHousing$Heating.QC) <- c("1", "2", "3", "4")
# levels(AmesHousing$Exterior.1st) <- c("Asbestos Shingles", "Brick Face", "Cement Board", "Hard Board","Metal Siding","Plywood","Stucco","Vinyl Siding","Wood Siding","Wood Shingles")
# levels(AmesHousing$"Foundation") <- c("Brick & Tile", "Cinder Block", "Poured Concrete")
# #Convert to Double
# AmesHousing$Garage.Qual <- as.numeric(as.character(AmesHousing$Garage.Qual))
# AmesHousing$Bsmt.Qual <- as.numeric(as.character(AmesHousing$Bsmt.Qual))
# AmesHousing$Heating.QC <- as.numeric(as.character(AmesHousing$Heating.QC))
```

```{r Recoding Variables, message=FALSE, warning=FALSE, echo = FALSE}
AmesHousing$home_type <- recode(AmesHousing$home_type, 
       `020` = "1-STORY 1946 & NEWER ALL STYLES", 
       `030` = "1-STORY 1945 & OLDER", 
       `040` = "1-STORY W/FINISHED ATTIC ALL AGES", 
       `045` = "1-1/2 STORY - UNFINISHED ALL AGES", 
       `050` = "1-1/2 STORY FINISHED ALL AGES", 
       `060` = "2-STORY 1946 & NEWER", 
       `070` = "2-STORY 1945 & OLDER",
       `075` = "2-1/2 STORY ALL AGES",
       `080` = "SPLIT OR MULTI-LEVEL", 
       `085` = "SPLIT FOYER", 
       `090` = "DUPLEX - ALL STYLES AND AGES", 
       `120` = "1-STORY PUD - 1946 & NEWER", 
       `150` = "1-1/2 STORY PUD - ALL AGES", 
       `160` = "2-STORY PUD - 1946 & NEWER",
       `180` = "PUD - MULTILEVEL - INCL SPLIT LEV/FOYER",
       `190` = "2 FAMILY CONVERSION - ALL STYLES AND AGES",
       .default = NULL)
AmesHousing$Fence <- recode(AmesHousing$Fence, 
            `GdPrv` =  "Good Privacy",
            `MnPrv`  =  "Minimum Privacy",
            `GdWo`   =  "GoodWood", 
            `MnWw`   = "Minimum Wood/Wire", 
            .default = NULL)

AmesHousing<-AmesHousing%>%
  mutate(Fence = ifelse(is.na(Fence), "No Fence", Fence))
AmesHousing$Land.Slope <- recode(AmesHousing$Land.Slope, 
            `Gtl` =  "Gentle Slope",
            `Mod`  =  "Moderate Slop",
            `Sev`   =  "Severe", 
            .default = NULL)
AmesHousing$Street <- recode(AmesHousing$Street, 
            `Grvl` =  "Gravel",
            `Pave`  =  "Paved",
            .default = NULL)
AmesHousing$Kitchen.Qual <- recode(AmesHousing$Kitchen.Qual, 
            `Ex` =  5,
            `Gd` =  4,
            `TA` =  3,
            `Fa` =  2,
            `Po` =  1, 
            .default = NULL)
AmesHousing$Neighborhood <- recode(AmesHousing$Neighborhood, 
       `Blmngtn` = "Bloomington Heights", 
       `Blueste` = "Bluestem", 
       `BrDale` = "Briardale", 
       `BrkSide` = "Brookside", 
       `ClearCr` = "Clear Creek", 
       `CollgCr` = "College Creek",
       `Crawfor` = "Crawford",
       `Edwards` = "Edwards",
       `Gilbert` = "Gilbert", 
       `IDOTRR` = "Iowa DOT and Rail Road", 
       `MeadowV` = "Meadow Village", 
       `Mitchel` = "Mitchell", 
       `Names` = "North Ames", 
       `NoRidge` = "Northridge",
       `NPkVill` = "Northpark Villa",
       `NridgHt` = "Northridge Heights",
       `NWAmes` = "Northwest Ames",
       `OldTown` = "Old Town",
       `SWISU` = "South & West Iowa State University",
       `Sawyer` = "Sawyer",
       `SawyerW` = "Sawyer West",
       `Somerst` = "Somerset",
       `StoneBr` = "Stone Brook",
       `Timber` = "Timberland",
       `Veenker` = "Veenker",
       .default = NULL)

AmesHousing$Garage.Qual<- recode(AmesHousing$Garage.Qual,
            `Ex` =  5,
            `Gd` = 4,
            `TA` = 3,
            `Fa` = 2,
            `Po` = 1,
            .default = NULL)

AmesHousing$Exterior.1st <- recode(AmesHousing$Exterior.1st,
       `AsbShng` = "Asbestos Shingles",
       `Wd Sdng` = "Asphalt Shingles",
       `BrkComm` = "Brick Common",
       `BrkFace` = "Brick Face",
       `CBlock` = "Cinder Black",
       `CemntBd` = "Cement Board",
       `HdBoard` = "Hard Board",
       `ImStucc` = "Imitation Stucco",
       `MetalSd` = "Metal Siding",
       `Other` = "Other",
       `Plywood` = "Plywood",
       `PreCast` = "PreCast",
       `Stone` = "Stone",
       `Stucco` = "Stucco",
       `VinylSd` = "Vinyl Siding",
       `Wd Sdng` = "Wood Siding",
       `WdShing` = "Wood Shingles",
       .default = NULL)

AmesHousing$Foundation <- recode(AmesHousing$Foundation,
       `Wood` = "Wood",
       `Stone` = "Stone",
       `Slab` = "Slab",
       `PConc` = "Poured Concrete",
       `CBlock` = "Cinder Block",
       `BrkTil` = "Brick & Tile",
       .default = NULL)

AmesHousing$Heating.QC <- recode(AmesHousing$Heating.QC,
            `Ex` =  5,
            `Gd`  = 4,
            `TA`   = 3,
            `Fa`   = 2,
            `Po` = 1,
            .default = NULL)

table(AmesHousing$Garage.Qual)
table(AmesHousing$Heating.QC)
table(AmesHousing$Foundation)
table(AmesHousing$Exterior.1st)
table(is.na(AmesHousing$Exterior.1st))
table(is.na(AmesHousing$Foundation))
table(is.na(AmesHousing$Heating.QC))
table(is.na(AmesHousing$Garage.Qual))


```



# Introduction 

The aim of this project....

This project used data from 1500 residential property sales in Ames, Iowa between 2006 and 2012. There are 82 explanatory variables in the data set, containing -  nominal, ordinal, discrete, and continuous attributes.  Continuous variables provide information about the multiple area dimensions of the house and property, such as the the size of the lot, garage among others. Discrete variables, on the other hand, quantify characteristics of the house/properties like the number of kitchens, baths, bedrooms, and parking spots. Nominal variables, generally, describe the multiple types of materials and locations, such name of the neighborhood or the type of foundations. Ordinal variables typically rate the condition and quality of multiple house characteristics  and utilities. 



# Exploratory Data Analysis 

We decided to keep this as a continuous variable as opposed to switching it to a factor. We did so because changing it to a factor would have lead to us dropping the "Very Poor" or "1" factor level as this level only has around 4 observations. By keeping the variable continuous, we are able to keep these observations and so better predict the home prices of homes that fall under this category.




### Exploring Selected Home Characteristics in the Dataset


Sale Price graph


```{r, warning = FALSE, echo = FALSE}
# ggplot(AmesHousing, aes(x = saleprice)) +
#   xlab("Sale Price (log)") +  geom_histogram()
# 
# ggplot(AmesHousing, aes(x = log(saleprice+1))) +
#   xlab("Sale Price (log)") +  geom_histogram()
```





```{r Counts by lot area, echo=FALSE, warning=FALSE, message=FALSE, out.width="800px", out.height="600px"}
count_lotA<-ggplot(AmesHousing, aes(x=lot_area)) + 
  geom_histogram(binwidth=500, fill="darkred") +
  theme(legend.position = "none") +
  xlab("Lot Area") +
  ylab("Count of Homes") +
  ggtitle("Figure 2: Count of Homes by Lot Area")
outlier.no <- length(boxplot.stats(AmesHousing$lot_area)$out)
outlier.sum <- summary(boxplot.stats(AmesHousing$lot_area)$out)
lot_area.outlier.rm <- filter(AmesHousing, lot_area < 17755)
count_lotA_2<- ggplot(lot_area.outlier.rm, aes(x=lot_area)) + 
  geom_histogram(binwidth=500, fill="darkred") +
  theme(legend.position = "none") +
  xlab("Lot Area") + 
  ggtitle(" ")
library(cowplot)
plot_grid(count_lotA, count_lotA_2)
#https://stackoverflow.com/questions/1249548/side-by-side-plots-with-ggplot2
```

```{r Counts by home type, message=FALSE, warning=FALSE, echo = FALSE}
AmesHousing <- within(AmesHousing, home_type <- factor(home_type, levels=names(sort(table(home_type), decreasing=TRUE))))
# https://stackoverflow.com/questions/5208679/order-bars-in-ggplot2-bar-graph
 ggplotly(ggplot(AmesHousing, aes(x=home_type)) + 
  geom_bar(stat="count", fill="darkred") +
  theme(axis.text.x=element_text(angle=-90, hjust=0)) +
  theme(legend.position = "none") +
  xlab("Type of Home") +
  ylab("Count") +
  ggtitle("Figure 3: Count of Home Type"))
```


```{r counts by year built, echo=FALSE, warning=FALSE, message=FALSE}
ggplotly(ggplot(AmesHousing, aes(x=year_built)) + 
  geom_histogram(binwidth=5, fill="darkred") +
  theme(legend.position = "none") +
  xlab("Year Built (5 Year Increments)") +
  ylab("Count of Homes") +
  ggtitle("Figure 4: Count of Homes by Year Built"))
```


When it comes to lot area, this dataset has many outliers as shown above. We found that there were `r outlier.no` outliers greater than the minimum outlier value of `r outlier.sum[1]`. As these made visualization difficult, we temporarily removed them.
After removing the outliers, we can see that homes have a somewhat normal distribution in terms of lot area near the median of `r summary(AmesHousing$lot_area)[3]` square feet.


From **Figure 3**, we see that 1-story homes that were built in 1946 or later make up the bulk of our dataset, specifically `r nrow(filter(AmesHousing, home_type=="1-STORY 1946 & NEWER ALL STYLES"))`. This is over one-third of our total dataset which has `r nrow(AmesHousing)` observations. Please not that the graphs are interactive so move your cursor over the graph to see more details.
Furthermore, we can also observe from **Figure 4**, that most homes were built within a 5 year time range of 2005. 



Summary Statistics
```{r, echo = FALSE}
# table(AmesHousing$Kitchen.Qual)
summary(AmesHousing$Kitchen.Qual)
```



## Relationship Between Sale Price and Selected Characteristics 


Some intro text 




```{r Saleprice distribution Neighborhood, message=FALSE, warning=FALSE, echo=FALSE, out.width="800px", out.height="800px"}
  
ggplotly(AmesHousing %>%
  ggplot( aes(x=reorder(Neighborhood, saleprice, FUN = median), y=saleprice)) +
    scale_fill_viridis(discrete = TRUE, alpha=0.5) +
    geom_jitter(color=rgb(0.1,0.4,0.5,0.7), size=0.4, alpha=0.5) +
   geom_boxplot(fill = rgb(0.1,0.4,0.5,0.7)) +
    theme(
      legend.position="none",
      plot.title = element_text(size=11)
    ) +
    ggtitle("Figure 5: Sale Price vs. Neighborhood") +
    xlab("Neighborhood") +
    ylab("Sale Price")+
    scale_y_continuous(labels = dollar)+
  theme(plot.title = element_text(hjust = 0.5),axis.text.x =  element_text(angle = -90)))
#reorder(Species, Sepal.Width, FUN = median)
```


```{r saleprice distribution x overal q, echo=FALSE, warning=FALSE, message=FALSE, out.width="800px", out.height="600px"}
graph6<- ggplot(AmesHousing, aes(overall_qual, saleprice)) + 
  geom_point(stat="identity", color=rgb(0.1,0.4,0.5,0.7)) +
  theme(legend.position = "none") +
  xlab("Overall Quality (10 is Best)") +
  ylab("Average Sale Price") +
  scale_y_continuous(labels = dollar)+
  ggtitle("Figure 6: Home Price By Overall Quality") +
  geom_smooth()
graph7<- ggplot(AmesHousing, aes(year_built, saleprice)) + 
  geom_point(stat="identity", color=rgb(0.1,0.4,0.5,0.7)) +
  theme(legend.position = "none") +
  xlab("Year Built") +
  ylab("Average Sale Price") +
  scale_y_continuous(labels = dollar)+
  ggtitle("Figure 7: Average Home Price By Year Built") +
  geom_smooth(method=lm)
plot_grid(graph6, graph7)
```



```{r Saleprice distribution home type, message=FALSE, warning=FALSE, echo=FALSE, out.width="800px", out.height="800px"}
 ggplotly(AmesHousing %>%
  ggplot(aes(fct_rev(fct_reorder(home_type, saleprice)), saleprice)) +
    scale_fill_viridis(discrete = TRUE, alpha=0.5) +
    geom_jitter(color=rgb(0.1,0.4,0.5,0.7), size=0.4, alpha=0.5) +
    geom_boxplot(fill = rgb(0.1,0.4,0.5,0.7)) +
    theme(legend.position="none", plot.title = element_text(size=11)) +
    ggtitle("Figure 8: Sale Price vs. Home Type") +
    xlab("Home Type") +
    ylab("Sale Price") +
    scale_y_continuous(labels = dollar)+
    theme(plot.title = element_text(hjust = 0.5), axis.text.x=element_text(angle=-90, hjust=0)))
```


```{r sale price distribution  kitchen qual, warning=FALSE, message= FALSE, echo=FALSE,out.width="800px", out.height="600px"}
graph9<- ggplot(AmesHousing, aes(Kitchen.Qual, saleprice)) +
  geom_point(stat="identity", color=rgb(0.1,0.4,0.5,0.7)) +
  theme(legend.position = "none") +
  xlab("Overall Quality (5 is Best)") +
  ylab("Sale Price") +
  scale_y_continuous(labels = dollar)+
  ggtitle("Figure 9: Home Price By Kitchen Quality") +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2), aes(colour = "Quadratic fit"))


graph10<-ggplot(lot_area.outlier.rm, aes(lot_area, saleprice)) + 
  geom_point(stat="identity", color=rgb(0.1,0.4,0.5,0.7)) +
  theme(legend.position = "none") +
  xlab("Lot Area") +
  ylab("Average Sale Price") +
  scale_y_continuous(labels = dollar)+
  ggtitle("Figure 10: Home Price By Lot Area") +
  geom_smooth(method=lm)
plot_grid(graph9, graph10)
```



**Neighborhood** We can observe a large variation in sale price across across different neighborhoods. Even within neighborhood we also see variation. Investigating some housing characteristics may give us insight into the variation observed in price within neighborhoods. 


In terms of overall quality, as expected price increases as overall quality increases. 


As one would expect, the newer a home is, the higher its price, on average.

When looking at home type by sale price, we find that 2 story homes built in the year 1946 or later have the highest median home prices.



We see a non-linear relationship between kitchen quality and sale price. The higher the kitchen quality the higher the median sale price. From **Figure 10**, we can see that - as expected - there is a gradual positive relationship between lot area and sales price.


## Hypothesis



# Methodology

```{r Splitting into test/train , message=FALSE, warning=FALSE, echo=FALSE}

AmesHousing_Short<-AmesHousing%>%
    select(
    full_bath_abv_grd,
    tot_rms_abv_grd,
    home_type,
    overall_qual,
    lot_area,
    year_built,
    Garage.Qual,
    Exterior.1st,
    Foundation,
    Bsmt.Qual,
    Heating.QC,
    Roof.Style,
    Kitchen.Qual,
    Neighborhood,
    Fence,
    Street,
    Land.Slope,
    Yr.Sold,
    saleprice)

training <- AmesHousing_Short%>%
  filter(Yr.Sold != 2010)

testing <- AmesHousing_Short%>%
  filter(Yr.Sold == 2010)


# Remove missing values
training<-training[complete.cases(training), ]
```


```{r forward/backward subset , message=FALSE, warning=FALSE, echo=FALSE}



#forward selection
regfit.fwd=regsubsets(saleprice~.,
                data=AmesHousing_Short,
                nbest = 1, 
                nvmax=19,
                method="forward")

#summary(regfit.fwd)

# BIC selected forward stepwise model
coef(regfit.fwd, which.min(summary(regfit.fwd)$bic))
plot(regfit.fwd,scale="bic")
#backward Selection

regfit.bwd=regsubsets(saleprice~.,data=AmesHousing_Short,nvmax=19,method="backward")


# BIC selected backward stepwise model
coef(regfit.bwd, which.min(summary(regfit.bwd)$bic))
plot(regfit.bwd,scale="bic")


```






*****SHANNON'S SUBSET********
```{r}

```


```{r}

```



*****JAY'S SUBSET********
```{r}

```


```{r}

```

```{r}

```




# Analysis


# Discussion


# References


```{r Remove Small N categories, message=FALSE, warning=FALSE, echo = FALSE}
AmesHousing <- AmesHousing %>% 
  filter(home_type != "1-STORY W/FINISHED ATTIC ALL AGES" & home_type != "1-1/2 STORY PUD - ALL AGES")%>%
  filter(!Neighborhood%in%c("GrnHill", "Landmrk", "Greens"))%>%
  filter(Kitchen.Qual!=1)
```
